---
title: 'Seminar 10: Descriptive statistics'
output: html_notebook
---

## 1. Data (Id is the average of the member's no. in the P&S list)

### Task: Generate your data 

This is a sample of size $n=1000$ combined from two samples of size $500$ from the normal distributions $N(\mu_k,\sigma_k^2)$ with $\mu_1 = Id$, $\sigma_1 = 3$ and $\mu_2 = -2$, $\sigma_2 = Id$

Hint: you can use $\texttt{mean = c}(\mu_1,\mu_2)$ and $\texttt{sd = c}(\sigma_1,\sigma_2)$

```{r}
library(moments)
```

```{r}
id <- 30

# your code here
id <- 50
n <- 1000

mu1 <- id
mu2 <- -2
sigma1 <- 3
sigma2 <- id

data <- rnorm(n, mean = c(mu1, mu2), sd = c(sigma1, sigma2))

summary(data)
``` 

## 2. Visualise the data

### Task: Draw the histogram of the data, empirical density, and empirical cdf. Comment on whether the data are close to a normal one

```{r}
# draw histogram
hist(data, breaks = 100, prob = TRUE)

# superimpose empirical density
lines(density(data), col = "red", lwd = 2)

# superimpose normal density
lines(density(rnorm(n, mean = mean(data), sd = sd(data))), col = "blue", lwd = 2)
```

### Plot ecdf

```{r} 
# plot ecdf
data_cdf = ecdf(data) 
plot(data_cdf, lwd = 2)

# superimpose cdf for standard normal
theoretical_data <- rnorm(n, mean(data), sd(data))
theoretical_cdf <- ecdf(theoretical_data)
lines(theoretical_cdf, col = "red", lwd = 2)

# calculate maximal difference
theoretical_cdf(5)
# max(knots(data_cdf) - knots(theoretical_cdf))
```

## 3. Skewness and kurtosis of the data

**Skewness** is defined as
$$
  \mathsf{E} \left[\left({\frac {X-\mu }{\sigma }}\right)^{3}\right]={\frac {\mu _{3}}{\sigma ^{3}}}={\frac {\mathsf{E} \left[(X-\mu )^{3}\right]}{\ \ \ (\mathsf{E} \left[(X-\mu )^{2}\right])^{3/2}}}
$$

Positive skewness means the data have longer right tail and then the mean value $\mathsf{E}(X)$ is greater than the median

**Kurtosis** is defined as 
$$
  \mathsf{E} \left[\left({\frac {X-\mu }{\sigma }}\right)^{4}\right]={\frac {\mu _{4}}{\sigma ^{4}}}={\frac {\mathsf{E} \left[(X-\mu )^{4}\right]}{\ \ \ (\mathsf{E} \left[(X-\mu )^{2}\right])^{2}}}
$$

For a standard normal distribution, the kurtosis is $3$. If we get a larger value, then more data are concentrated ner the mean, and the empirical density is more steep


```{r}
library(moments)

# skewness
data_skewness <- skewness(data)
data_skewness
# kurtosis
data_kurtosis <- kurtosis(data)
data_kurtosis
```

## 4. Percentiles 

### For q= 0.1,...,0.9 calculate sample q-percentiles manually (i.e. by sorting the data and picking the corresponding values) and by using percentile

```{r} 
# your code here
data_sorted <- sort(data)
percentiles_handy <- c()
for (i in 1:9) {
  percentiles_handy[i] <- data_sorted[i*(n/10)]
}
print(percentiles_handy)

print(quantile(data_sorted, probs = c(1:9)/10))
```


## 5. Sample mean and sample standard error vs the theoretical ones 

### What is the theoretical expected value of the random variable considered? What is the variance? Compare them to the sample values; explain the difference in the variance

```{r}
# theoretical expected value and variance
theoretical_mu <- mu1 + mu2
theoretical_sigma <- sqrt(sigma1^2 + sigma2^2)

# sample mean and sample variance
```


## 6. k-sigma rule

### Calculate the fraction of the data that are within $k\sigma$ of the sample mean for $k=1,2,3$. Do we get the result expected? Why or why not?

```{r}
# your code here
```

